A Systematic Review of the Use of the ESM Methodology in Education

### 2020-09-16 paper_draft.Rmd
last 2 sections for missing data and analysis need to be finished.

### Sample Size
The sample size was investigated to determine how many participants were in each of the studies. The sample size ranged from 1 participant to a study with over one thousand student participants. Determining and coding for sample size was difficult depending on the article. Some studies would refer to their sample size as the number of participants, yet they ran their analysis with less than the originally stated value. During inter-rator training, we determined that the sample size would be the number of individuals that were analyzed and contributed to the data. 
### Prompting Method
For a study to be truly and ESM study, the participants needed to receive some form of prompt that would signal the participant to contribute. The prompting method refers to when and what style that participants were signaled. For example, in some studies participants were signaled at random times during a given period, this could either be during a class period, a lab, or even random times during an entire day. Other studies used set time as their prompting method. Participants would be signaled at specific times during the day. Then others would use the participants' schedule as the frame for a set time that they would answer survey questions, such as their bedtime or during study time for their courses. 
### Number of Signals
Signals affect the power of the study. It is said the more signals the better, this depends on what you are investigating and how long you have to run the study. During coding determining the number of signals each of the participants received was not stated overtly. On occasion, we would have to take bits of information that were provided by the authors and mathematically estimate the number of signals that the participants received. Some of the information that was provided was the number of signals that were sent in total and then we would have to divide that by the number of participants. This gave us a decent estimate of the number of signals sent. This became difficult when articles that stated the number of participants were different throughout the paper. 
Others would not even provide how many times each participant received signals to participate in the ESM study. We, the coders, could not extrapolate the number of signals from the information that was provided to us. Then finally, there were some articles where the authors provided exact numbers of how many signals each person received. 
We had difficulty when authors would state the number of signals per day, but did not provide a definite number of days, weeks, or months for the ESM study. These papers had such limited facts of length that we could not calculate the number of signals that participants received. 
### Length of Signaling Period (study length)
This is a determining factor for determining how many signals a participant received. The longer the signaling period tended to equate to more signals received by the individual participants. The signal length varied widely, from 1 lesson on a single day to 7 weeks or another that was vague and stated it occurred over a whole semester. 
Coding for the signaling period was difficult at times. Some of the authors overtly stated the length of the period, while others did not give any indication of how long the ESM was used. This also caused a problem when determining the number of signals that students received. 
### Response Rate
During the coding of the review papers, this became something that we would continually look for in a paper. Many papers (n=  ) did not provide the reader with the response rate for their study. This takes away from the power of the study. There were other times we were required to calculate the response rate. We would use the number received and divide by the number of participants. This was difficult at times when the number of participants was unclear. 
### Signal Method
How were participants informed when it was time for them to fill out the survey/questionnaires.
ESM has evolved in the methods that the participants are signaled when to fill out the surveys, questionnaires, or other documents. Signaling methods were sometimes not provided, but other factors in the paper such as the authors stating that it was an ESM study allowed us the justification to keep the study in this review. 
Signaling methods were as simple as a pencil or pen and a piece of paper done in the classroom setting signaled by the teacher. Other we combined in a packet that was taken home and combined with a pager or other electronic device. 
Devices that ESM researchers used included: pager, cell phone (later transformed to “app”), also apps that could be used on tablets (Android, iPad, etc.), PDA, and text messages. Some of the originals coded for signaling methods were transformed and evolved as we discussed coding and came to a consensus. For example, we looked at the usage of cell phones as one single code, but later we separated into apps for those that were not a form of text messages and text messages. Then text messages were further described, coded, as SMS since the ESM programs usually use an SMS format for sending messages to participants. 
### Items used
We also focused on the designs of the investigation, what type of question models researchers used. For example, each of the items that researchers used was instruments to determine what each participant was experiencing. Researchers many times used a type of Likert scale, psychometric scale, which varied from a 3 point Likert scale up to a 100-point Likert scale. This varied according to the needs and the objectives of the researchers.  Then researchers would use open-ended questions to enable participants to elaborate on their experiences. Finally, some authors were unclear of what instruments were used. They would only state that they used a questionnaire and no further information was provided. 
### Description of items
To further draw information and specifics of the ESM that was used in research we looked at what types of topics the items were seeking to glean data from the participants. We wanted to know what kinds of question topics that have been used in the current literature. Some articles would only elaborate and say that more than one Likert scale was used, while others would go into detail about the different topics. For example, in the Shumow et al. (2008), they had 3 items on the cognitive state, 4 on affective, 2 on the motivational state, and 2 indication student self-views.  Detail of what items were used it students varied from detailed to no information provided, minus what kind of items, instruments, were used in their study. 
### Number of Items
We wanted to know the number of items that were used in each of the articles. Some would only use a single item to determine the experience of the participants, while others would use up to 86 items in their surveys. 
### Validity and Reliability
For Validity and Reliability, we focused on the authors providing a Cronbach’s alpha or not. Though there are other methods to determine validity and reliability, an alpha value is commonly used and effective when determining the significance of using a Likert scale questionnaire. At the beginning of coding, we discussed other methods for Validity and Reliability, such as P-value and others, but we came to a consensus that alpha would be the most functional and over encompassing for this review with the number of articles that used Likert scales as their instruments.
### Missing Data Strategies
Researchers accounting for why certain collected data were not analyzed. 
Though many articles did not provide the reader an accounting for removal or lack of data. Some articles did give and account for missing data by using, for example, Full Information Maximum Likelihood (FIML), listwise deletion, or pairwise elimination.

>>>>

### Sample Size
The sample size was investigated to determine how many participants were in each of the studies. The sample size ranged from 1 participant to a study with over one thousand student participants. Determining and coding for sample size was difficult depending on the article. Some studies would refer to their sample size as the number of participants, yet they ran their analysis with less than the originally stated value. During inter-rator training, we determined that the sample size would be the number of individuals that were analyzed and contributed to the data.

### Prompting Method
For a study to be truly and ESM study, the participants needed to receive some form of prompt that would signal the participant to contribute. The prompting method refers to when and what style that participants were signaled. For example, in some studies participants were signaled at random times during a given period, this could either be during a class period, a lab, or even random times during an entire day. Other studies used set time as their prompting method. Participants would be signaled at specific times during the day. Then others would use the participants' schedule as the frame for a set time that they would answer survey questions, such as their bedtime or during study time for their courses.

### Number of Signals
Signals affect the power of the study. It is said the more signals the better, this depends on what you are investigating and how long you have to run the study. During coding determining the number of signals each of the participants received was not stated overtly. On occasion, we would have to take bits of information that were provided by the authors and mathematically estimate the number of signals that the participants received. Some of the information that was provided was the number of signals that were sent in total and then we would have to divide that by the number of participants. This gave us a decent estimate of the number of signals sent. This became difficult when articles that stated the number of participants were different throughout the paper. 
Others would not even provide how many times each participant received signals to participate in the ESM study. We, the coders, could not extrapolate the number of signals from the information that was provided to us. Then finally, there were some articles where the authors provided exact numbers of how many signals each person received. 
We had difficulty when authors would state the number of signals per day, but did not provide a definite number of days, weeks, or months for the ESM study. These papers had such limited facts of length that we could not calculate the number of signals that participants received.

### Length of Signaling Period (study length)
This is a determining factor for determining how many signals a participant received. The longer the signaling period tended to equate to more signals received by the individual participants. The signal length varied widely, from 1 lesson on a single day to 7 weeks or another that was vague and stated it occurred over a whole semester. 
Coding for the signaling period was difficult at times. Some of the authors overtly stated the length of the period, while others did not give any indication of how long the ESM was used. This also caused a problem when determining the number of signals that students received.

### Response Rate
During the coding of the review papers, this became something that we would continually look for in a paper. Many papers (n=  ) did not provide the reader with the response rate for their study. This takes away from the power of the study. There were other times we were required to calculate the response rate. We would use the number received and divide by the number of participants. This was difficult at times when the number of participants was unclear.

### Signal Method
How were participants informed when it was time for them to fill out the survey/questionnaires.
ESM has evolved in the methods that the participants are signaled when to fill out the surveys, questionnaires, or other documents. Signaling methods were sometimes not provided, but other factors in the paper such as the authors stating that it was an ESM study allowed us the justification to keep the study in this review. 
Signaling methods were as simple as a pencil or pen and a piece of paper done in the classroom setting signaled by the teacher. Other we combined in a packet that was taken home and combined with a pager or other electronic device. 
Devices that ESM researchers used included: pager, cell phone (later transformed to “app”), also apps that could be used on tablets (Android, iPad, etc.), PDA, and text messages. Some of the originals coded for signaling methods were transformed and evolved as we discussed coding and came to a consensus. For example, we looked at the usage of cell phones as one single code, but later we separated into apps for those that were not a form of text messages and text messages. Then text messages were further described, coded, as SMS since the ESM programs usually use an SMS format for sending messages to participants.

### Items used
We also focused on the designs of the investigation, what type of question models researchers used. For example, each of the items that researchers used was instruments to determine what each participant was experiencing. Researchers many times used a type of Likert scale, psychometric scale, which varied from a 3 point Likert scale up to a 100-point Likert scale. This varied according to the needs and the objectives of the researchers.  Then researchers would use open-ended questions to enable participants to elaborate on their experiences. Finally, some authors were unclear of what instruments were used. They would only state that they used a questionnaire and no further information was provided. 
### Description of items
To further draw information and specifics of the ESM that was used in research we looked at what types of topics the items were seeking to glean data from the participants. We wanted to know what kinds of question topics that have been used in the current literature. Some articles would only elaborate and say that more than one Likert scale was used, while others would go into detail about the different topics. For example, in the Shumow et al. (2008), they had 3 items on the cognitive state, 4 on affective, 2 on the motivational state, and 2 indication student self-views.  Detail of what items were used it students varied from detailed to no information provided, minus what kind of items, instruments, were used in their study.

### Number of Items
We wanted to know the number of items that were used in each of the articles. Some would only use a single item to determine the experience of the participants, while others would use up to 86 items in their surveys.

### Validity and Reliability
For Validity and Reliability, we focused on the authors providing a Cronbach’s alpha or not. Though there are other methods to determine validity and reliability, an alpha value is commonly used and effective when determining the significance of using a Likert scale questionnaire. At the beginning of coding, we discussed other methods for Validity and Reliability, such as P-value and others, but we came to a consensus that alpha would be the most functional and over encompassing for this review with the number of articles that used Likert scales as their instruments.

### Missing Data Strategies
Researchers accounting for why certain collected data were not analyzed. 
Though many articles did not provide the reader an accounting for removal or lack of data. Some articles did give and account for missing data by using, for example, Full Information Maximum Likelihood (FIML), listwise deletion, or pairwise elimination.
